%%% -*-LaTeX-*-

\chapter{Introduction}
\label{cha:introduction}

% A few sentences on the priming project

Interactions between the blood and artificial materials is an
important engineering problem. Artificial devices are commonly used to
treat cardiovascular diseases, and they must be designed to not
trigger immune responses in the bloodstream. However, current designs
are not sufficiently blood-compatible and patients with a
blood-contacting device must be placed on systemic anticoagulants to
reduce the risks of thrombosis and embolism
\cite{Ratner1993,Ratner2007,Oprea13}. 
	
% 2. Contextualize rolling within blood clotting
% Process of clotting begins when agonists (vWF and collagen) are
% exposed on the ECs

\section{Platelet rolling within the clotting pathway}
\label{sec:plat-roll-with}

There are three major processes in the blood clotting pathway:
platelet activation and adhesion, coagulation, and fibrin
gelation. In platelet adhesion and activation, chemical platelet
agonists are exposed on the vessel wall and released into the fluid,
signaling platelets to begin clot formation and adhere to the vessel
wall. Once activated platelets begin to aggregate, extracellular
chemical reactions begin to occur both in the blood and on platelet
surfaces ultimately resulting in the production of thrombin. This set
of extracellular reactions is referred to as coagulation.

The process of platelet activation is mediated through
intracellular chemical pathways. When a platelet binds with agonists,
activation pathways are triggered which result in a suite of
activation responses: degranulation (i.e. release of ADP and other
soluble agonists), thromboxane A2 (TxA2) synthesis, activation and
recruitment of \ITA{IIb}\ITA{3} receptors, exposure of
phosphatidylserine (PS), and---after adhesion to a vessel
wall---platelet spreading. (Need to source this)

Coagulation broadly refers to the extracellular chemical reaction
network that produces thrombin


\subsection{Platelet margination}
\label{sec:platelet-margination}

In order for a platelet aggregate to form, two processes need to
occur: platelet adhesion to the vessel wall, and cohesion among
platelets. In normal hemostasis, adhesion is facilitated by two
surface-immobilized proteins: collagen embedded in the sub-endothelial
matrix, and von Willebrand factor (vWF) which can either be bound to
collagen in the SE matrix or expressed by endothelial cells in
pathological circumstances (see refs in Fogelson and Neeves, 2015,
pp. 384 \& 390).
		
% Platelets bind to these agonists with fast receptors first (GP1b and
% GPVI), and then later with slow receptors (\alpha{IIb} \beta{3} and
% \alpha{2} \beta{1}).
Platelets constitutively express GP1b and GPVI receptors for vWF and
collagen respectively (see refs in Fogelson and Neeves 2015,
p. 384). These receptors have fast binding and unbinding kinetics, and
are therefore able to form bonds with their surface-immobilized
ligands even when the platelets are in flows with high wall shear
rates. When these bonds form and are stretched beyond their rest
length, they exert a force in opposition to the flow and slow the
platelet down. Therefore the motion of a platelet along a wall with
immobilized agonists is a function of the biochemical interactions of
receptors and ligands as well as the fluid forces exerted on the
platelet.
		
% The fast receptors trigger intracellular signaling cascades that
% result in platelet activation and activation of the slow receptors.
These receptors are more than simple physical links between the
platelet and surface. When they are bound with a ligand, they initiate
intracellular signaling cascades that are responsible for triggering
release of intracellular calcium and activating
phosphatidylinositide-3-kinase (PI3K) which are two crucial events in
platelet activation (Bye paper; look at Fogelson and Neeves; Du
paper?). These activation pathways ultimately terminate in a suite of
responses that are collectively called "platelet activation,"
including granule secretion, TxA2 synthesis, cytoskeletal
rearrangements, and activation of integrins \ITA{IIb}\ITB{3} and
\ITA{2}\ITB{1}. These integrins are receptors for fibrinogen/vWF
and collagen, respectively. They are constitutively expressed on the
surface of the platelet (check this!), but on unactivated platelets
they exist mostly in their low-affinity conformation (sources?).
		
In general, integrins form a large group of transmembrane receptors
that are primarily involved in the adhesion of cells to extracellular
matrix (ECM) and cell locomotion. They have a large extracellular
domain called the ectodomain with a hinge. In the low affinity
conformation, the ectodomain is bent at the hinge and the ligand
binding domain is at least partially blocked. When switching from
low-affinity to high-affinity conformations the integrin ectodomain
extends at the hinge to reveal the ligand binding site (sources). Note
I've avoided using the words "inactive" and "active." This is
intentional: there is evidence that integrins retain some ability to
form bonds with their ligands when in the low-affinity conformation
(sources), and experiments showing that resting platelets can bind to
fibrinogen (sources) also suggest this is possible.
		
Once integrins are bound and in their high-affinity conformation, they
can activate signal transduction pathways to initiate the formation of
clusters of integrins that mediate firm adhesion of a cell to ECM
(sources), along with other responses.
	
% Once platelets are activated, they can bind firmly to the wall,
% release soluble platelet agonists which causes a platelet aggregate to
% form, and release thrombin which results in the formation of a fibrin
% gel which ultimately stabilizes a clot.

% 3. Description of cell rolling, a single cell interacting with the surface
% Should mention margination in here somewhere.
% Cell rolling is the important first part of the process of platelet
% activation. There are three main insoluble platelet agonists: vWF,
% collagen, and fibrinogen. Two important constitutively active platelet
% receptors are GP1b (binds to vWF) and GPVI (binds to collagen). These
% receptors have fast association and dissociation and mediate platelet
% rolling along a vessel wall.
% There are also constitutively expressed integrins \alpha{IIb}
% \beta{3} and \alpha{2} \beta{1}, however they are in their
% low-affinity conformation.

% 1. Description of the priming project

% Why is hemocompatibility important?

Cardiovascular disease is one of the leading causes of death in the
U.S, and a common treatment for these diseases is to implant medical
devices into the blood stream. For example, stents (an expandable
solid mesh) are often used to treat stenotic arteries. However, the
introduction of a foreign material into the blood stream will cause
thrombosis unless the material is treated somehow, and while these
devices have been effective in saving lives, they are still far from
perfect. Patients with these implanted devices still must be placed on
anticoagulants, as they have a higher risk of a thrombotic event even
when the device is functional and surgery is successful (cit: ref 9 in
Colin's thesis--Cannegieter et. al., 1994).
		
% What has been done so far in hemocompatibility research, and what is
% Vlado's group trying to add?

Most hemocompatibility studies focus on local interactions and effects
of the material on platelets. However recent work (cit: Dr. Hlady's
work) has shown that nonlocal effects are also important in
understanding platelet interactions with implanted biomaterials. In
particular they have shown that platelet interactions with immobilized
agonists may not cause the platelets to adhere at the site of
interaction, but can nonetheless prime them for downstream adhesion
and full activation. One possible example of this in a vascular graft
is shown in Figure \# something. At either end of the implanted
device, native tissue must be joined with the artificial device. At
these points along the vessel wall the tissue is inflamed and
anastomotic (source?), which could expose platelet agonists on the
surface of the wall. Additionally the increased shear rate in the
stenotic regions could act as a platelet agonist (source?). Therefore
while platelets may not bind to a single inflamed region of the
vessel, the upstream region may prime platelets for adhesion, and then
more readily bind to the downstream inflamed region. It is also
possible that the inflamed tissue or the biomaterial may not cause
platelets to bind locally, but nonetheless partially activate
platelets and cause them to adhere downstream somewhere in the
circulation.
		
% Describe experimental setup, ongoing experiments, etc
In order to investigate nonlocal effects of priming, Dr. Hlady's group
has designed a microfluidic assay with two regions printed with
platelet agonists (see Figure whatever). The upstream agonist region
will be called the priming region, and the downstream region with
agonist will be called the capture region. In previous research,
Dr. Hlady's group has found that all 3 immobilized platelet agonists
tested (vWF, collagen, and fibrinogen) in the upstream region
increased adhesion of platelets in the capture region relative to a
nonreactive control (Colin's thesis, Corum paper). Also, recent
unpublished work has shown similar results when the upstream agonist
printing is replaced by a stenotic region with higher wall shear rates
(reference Shek's work).

\section{Existing cell rolling models}
\label{sec:exist-cell-roll}

A large body of work has been published on the subject of cell rolling
and adhesion, including mathematical models of cell rolling. However
most of the modeling work has been done on leukocyte rolling on an
agonist-coated surface, and relatively less work has been done on
platelet rolling.

Adhesive dynamics (AD) is perhaps the most well-known modeling
framework for leukocytes, and the first AD model was published in 1992
by Hammer \& Apte \cite{Hammer1992}. The Hammer and Apte paper
provides a good example to describe the AD models in general. It is
convenient to view the more recent models as modifications of this
fundamental model. They model a leukocyte as a rigid sphere, with
rigid microvilli protruding normally from the surface of the
cell. Receptors are located only at the tips of these microvilli
(multiple receptors can occupy a single microvillus), and can bind to
ligand binding sites which are uniformly distributed along a flat
wall. The positions and bound state of each receptor is tracked
throughout the simulation. Binding and unbinding of receptors is
modeled with a stochastic process with length-dependent rates given by
the Dembo model \cite{Dembo1988}:
\begin{align}
  \label{eq:dembo-on}
  k_\tn{on} \left(L_\tn{sep}\right)
  &= k_\tn{on}^0 \exp \left( -\frac{\sigma_\tn{ts} (L_\tn{sep} -
    \lambda)^2}{2\boltzmann \temp} \right) \\
  \label{eq:dembo-off}
  k_\tn{off} \left(L_\tn{sep}\right)
  &= k_\tn{off}^0 \exp \left( \frac{(\sigma - \sigma_\tn{ts})
    (L_\tn{sep} - \lambda)^2}{2 \boltzmann \temp} \right).
\end{align}
where $\lambda$ is the equilibrium separation distance,
$\boltzmann\temp$ is the thermal energy, $\sigma$ is the spring
constant, and $\sigma_\tn{ts}$ is the transition state spring
constant.

The movement of the leukocyte is determined by balancing hydrodynamic
forces and various chemical forces. The hydrodynamic forces are found
by assuming that the fluid surrounding the leukocyte is a Newtonian
fluid with $\Reynolds = 0$ at the revelent length and time scales of
the problem, and the fluid is moving in a shear flow. Under this set
of assumptions, there is a linear relationship between the drag force
$\forceVec$ on the cell and the velocity $\velVec$: $\forceVec =
\resMatrix \velVec$. The chemical forces on the cell are split up into
5 different forces: bond forces between the cell and the surface, van
der Waals forces, gravitational body force, electrostatic force, and
steric stabilization force.
